---
title: "DRY coding"
author: "Data Science in Biomedicine - Week 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```
## Introduction

We'll be using an in-built data set from the MASS package. But, because the MASS package can be a bit problematic to install, I've provided the data for you. It's in .rds format, so we load it in using the `read_rds()` function.

For more info about the dataset, head to https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/survey.html

```{r}
library(tidyverse)
library(knitr)

survey <- read_rds("./data/survey.rds")
```

## Exercise 1
Explore the data - what variables are in there, what variable types, how many variables and observations?
```{r}
head(survey)
```


## Exercise 2
Round every numeric variable to 0 decimal places.

```{r}
survey2 <- survey %>% 
  mutate(across(
    c(Wr.Hnd, NW.Hnd, Height, Age), ~ round(.x, digits = 0)
  ))
```


## Exercise 3
Using the rounded dataframe you created in Exercise 2, calculate the mean of each numeric variable in the dataset. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
survey2 %>% 
  select(where(is.numeric)) %>%
  map_df(~mean(.x, na.rm = TRUE)) %>% 
    kable()
```


## Exercise 4
Now make exactly the same table, using the pivot functions. Again, use kable() to insert a nicely formatted table into your R Markdown document.
```{r}
survey3 <- survey2 %>% 
  select(where(is.numeric)) %>%
  pivot_longer(
    names_to = "Variables",
    cols = c(Wr.Hnd, NW.Hnd, Pulse, Height, Age),
    values_to = "Values" ) %>% 
  group_by(Variables) %>% 
  summarise(Mean = mean(Values, na.rm = TRUE))  %>% 
  pivot_wider(
    names_from = Variables,
    values_from = Mean
  )

survey3 %>% 
  kable()
```

## Exercise 5
Students had initially expressed their height in either imperial or metric units, but someone converted everything into centimeters. Create a new variable, Height_original, which will show height in centimeters for students who expressed their height in metric units, and height in inches for students who prefer imperial units. 

NOTE: This one is a bit of a challenge. You may want to check out the `case_when()` function.

```{r}
survey2 <- survey2 %>% 
  mutate(Height_original = case_when(
    M.I == "Metric" ~ Height,
    M.I == "Imperial" ~  Height/2.54
  ))
```


## Exercise 6
Create a table that shows the mean and standard deviation of height - in centimeters for students who used metric units, and in inches for students who used imperial units. Use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
survey4 <- survey2 %>%
  filter(is.na(M.I) != TRUE) %>% 
  group_by(M.I) %>%
  summarise(mean = mean(Height_original),
         sd = sd(Height_original))

survey4 %>% 
  kable()
```


## Exercise 7
Create a contingency table to see the relationship between exercise and smoking. Pivot your table to make it easier to read, and then use kable() to insert a nicely formatted table into your R Markdown document.

```{r}
survey5 <- survey2 %>% 
  group_by(Exer, Smoke) %>% 
  tally()

survey5 <- survey5 %>% 
  pivot_wider(
    names_from = Exer,
    values_from = n
  )
```


## Exercise 8
Create a grid of Pulse histograms - one histogram for each combination of Sex and Exercise.

```{r}
survey2 %>% 
  select(Pulse, Sex, Exer) %>% 
  ggplot(aes(x = Pulse)) +
  geom_histogram(bins = 10) +
  facet_grid(Sex~Exer)
```


## Exercise 9
Create a new variable, Hnd.diff, that shows the difference in span between the writing and non-writing hand. Create a set of boxplots of this variable, with a separate boxplot for those who write with their right hand, and another one for those who write with their left hand.

```{r}
survey2 <- survey2 %>% 
  mutate(Hnd.diff = Wr.Hnd - NW.Hnd)

survey2 %>% 
  ggplot(aes(x = Hnd.diff)) +
  geom_boxplot() +
  facet_wrap(~W.Hnd)
```

